---
title: "Supplementary File"
author: "APB"
date: "1/8/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ape)
library(plyr)
library(tidyverse)
library(ouch)
library(dplyr)
library(phytools)
library(geiger)
library(caper)
library(picante)
library(pmc)
library(corHMM)
library(OUwie)

```



Supplmental Figure 1. 249 taxa helminth molecular phylogeny

```{r load_in_APBs_helminth_tree, echo=FALSE}


## load tree
tree<-read.nexus("hel2_44burn_copy.trees")

## load the binomial names for the coded tree names
nom <- read.csv("name_key.csv")

# make nom characters not factors
nom$CorrectName<-as.character(nom$CorrectName)
nom$Code<-as.character(nom$Code)

## ensure that every tip has a binomial name
#which(!(tree$tip.label %in% nom[,2])) ## ZTGLUN does not have a binomial name

## drop it from the tree
hel_tree <- drop.tip(tree, "ZTGLUN")
tree <- drop.tip(hel_tree, "HABMUS") #nematodes in the wrong area of the tree
tree <- drop.tip(tree, "ASCLUM") #nematodes in the wrong area of the tree
tree <- drop.tip(tree, "ASCSUU") #nematodes in the wrong area of the tree
tree <- drop.tip(tree, "PELSTR") #nematodes in the wrong area of the tree

##binomial names for only the species in the tree
nom[nom$Code %in% tree$tip.label,]->a

##nom in same order as tree tip labels
b<- a[match(tree$tip.label, a$Code),]

#replace code names with binomial names
tree$tip.label<-b$CorrectName

#check
#plot(tree, cex=0.3, type = "f")

#rename the global tree for GMPD data
global_tree<-tree


global_tree<-drop.tip(global_tree, c("Filaria martis","Prosthenorchis elegans"))

plotTree(global_tree, fsize =0.3, type = "fan")
```












\newpage

```{r load_nearctic_data, include = FALSE}
#host-parasite association data
nearctic_assoc_data<-read.csv("Host_Range_Nearctic_Mammals_list.csv", header=T)

#change to characters instead of factors
nearctic_assoc_data$Host<-as.character(nearctic_assoc_data$Host)
nearctic_assoc_data$Parasite<-as.character(nearctic_assoc_data$Parasite)

#parasite abundance/occurence data
nearctic_abund_data<-read.csv("Host_Range_Nearctic_Mammals.csv")

#change to characters instead of factors
nearctic_abund_data$Parasite<-as.character(nearctic_abund_data$Var1)

## Read in GMPD data
GMPDmain<-read.csv("GMPD_main.csv")
GMPDparasite<-read.csv("GMPD_parasite_traits.csv")

## Load GMPD data and subset down to the species in the parasite phylogeny
GMPDsmall<-GMPDmain[GMPDmain$ParasiteCorrectedName %in% global_tree$tip.label, ]


##load complete mammal tree from PHYLACINE
large_mam_tree<-read.nexus("mammal_phylo_consensus.nex")


large_mam_tree$tip.label<-gsub("_", " ", large_mam_tree$tip.label)

setdiff(GMPDsmall$HostCorrectedName, large_mam_tree$tip.label)

#### Make sure host names in the GMPD match host names in the PHYLACINE tree ####


#Make names match Elton binomial names
GMPDsmall$HostCorrectedName<-gsub("Cercopithecus albogularis", "Cercopithecus mitis", GMPDsmall$HostCorrectedName)

GMPDsmall$HostCorrectedName<-gsub("Equus burchelli", "Equus quagga", GMPDsmall$HostCorrectedName)

GMPDsmall$HostCorrectedName<-gsub("Equus quaggai", "Equus quagga", GMPDsmall$HostCorrectedName)

GMPDsmall$HostCorrectedName<-gsub("Otaria flavescens", "Otaria byronia", GMPDsmall$HostCorrectedName)

GMPDsmall$HostCorrectedName<-gsub("Cercopithecus lhoesti","Allochrocebus lhoesti", GMPDsmall$HostCorrectedName)

GMPDsmall$HostCorrectedName<-gsub("Equus caballus","Equus ferus", GMPDsmall$HostCorrectedName)

GMPDsmall$HostCorrectedName<-gsub("Lama glama","Lama guanicoe", GMPDsmall$HostCorrectedName)

GMPDsmall$HostCorrectedName<-gsub("Bos frontalis","Bos gaurus", GMPDsmall$HostCorrectedName)

GMPDsmall$HostCorrectedName<-gsub("Equus asinus", "Equus hemionus", GMPDsmall$HostCorrectedName)

GMPDsmall$HostCorrectedName<-gsub("Equus asinus", "Equus hemionus", GMPDsmall$HostCorrectedName)

GMPDsmall$HostCorrectedName<-gsub("Monachus schauinslandi", "Neomonachus schauinslandi", GMPDsmall$HostCorrectedName)

GMPDsmall$HostCorrectedName<-gsub("Neotragus moschatus", "Neotragus pygmaeus", GMPDsmall$HostCorrectedName)

GMPDsmall$HostCorrectedName<-gsub("Ovis aries", "Ovis orientalis", GMPDsmall$HostCorrectedName)


##remove the ones without a name
GMPDsmall <- GMPDsmall[GMPDsmall$HostCorrectedName!="no binomial name",]


#### MAKE A MAMMAL TREE FOR HOSTS IN THE NEARCTIC DATASET ####

##prune large tree to only include mammals from the Nearctic dataset
setdiff(large_mam_tree$tip.label, nearctic_assoc_data$Host)->bye
nearctic_mammal_tree<-drop.tip(large_mam_tree, large_mam_tree$tip.label[which(large_mam_tree$tip.label%in%bye)])
plot(nearctic_mammal_tree, cex=0.3)

#check if any hosts are missing
setdiff(nearctic_assoc_data$Host, nearctic_mammal_tree$tip.label)->missing
missing # "Mustela vison", "Alces americanus", "Alopex lagopus", "Erethizon dorsatus"


#### MAKE A MAMMAL TREE FOR HOSTS IN THE GMPD DATASET ####

#prune small tree to only include mammals from the GMPD
setdiff(large_mam_tree$tip.label, GMPDsmall$HostCorrectedName)->bye2
GMPD_mammal_tree<-drop.tip(large_mam_tree, large_mam_tree$tip.label[which(large_mam_tree$tip.label%in%bye2)])
plot(GMPD_mammal_tree, cex=0.3)


##subset global tree to nearctic data
setdiff(global_tree$tip.label, nearctic_assoc_data$Parasite)->bye3
nearctic_tree<-drop.tip(global_tree, global_tree$tip.label[which(global_tree$tip.label%in%bye3)])
plot(nearctic_tree, cex=0.3)

#interspecific distance matrix of parasites
nearctic_dis_para<-cophenetic.phylo(nearctic_tree) 

#intespecific distance matrix of mammals 
nearctic_dis_mam<-cophenetic.phylo(nearctic_mammal_tree) 

##make the pipeline to get the absense/presence of 
##each parasite (rows) that infect a given host (column)

nearctic_mamlist<-as.vector(nearctic_mammal_tree$tip.label) #list of mammals in tip label order

nearctic_para_list<-as.vector(nearctic_tree$tip.label) #list of parasites

#make a dataframe with host names as columns and parasite names as rows
nearctic_df <- data.frame(matrix(ncol=length(nearctic_mamlist),nrow=length(nearctic_para_list)), row.names = nearctic_para_list)
colnames(nearctic_df)<-nearctic_mamlist

# complete dataframe with association data, 
# where 0 = not found in host and 1 = found in host

for (i in 1:nrow(nearctic_df))
  nearctic_df[i,] <- (colnames(nearctic_df) %in% unique(subset(nearctic_assoc_data, Parasite==rownames(nearctic_df)[i])$Host)) %>% as.numeric

#save matrix
write.csv(nearctic_df,file = "association_matrix_neoartic.csv", row.names = TRUE, col.names = TRUE)


## Create MPD scores for every parasite using the Phylacine mammal tree 
nearctic.para.mpd <- mpd(nearctic_df, nearctic_dis_mam)
nearctic.para.mpd[is.na(nearctic.para.mpd)] <- 0 ## one host species have an MPD of 0
names(nearctic.para.mpd)<-nearctic_para_list


## Create MPD scores for every parasite using the Phylacine mammal tree 
nearctic.para.mpd.remove <- mpd(nearctic_df, nearctic_dis_mam)
names(nearctic.para.mpd.remove)<-nearctic_para_list
nearctic.para.mpd.remove<-nearctic.para.mpd.remove[!is.na(nearctic.para.mpd.remove)] ## one host species MPD removed



## Create MPD scores for every host using the helminth tree
nearctic.mam.mpd <- mpd(t(nearctic_df), nearctic_dis_para)
nearctic.mam.mpd[is.na(nearctic.mam.mpd)] <- 0 ## hosts infected by only one parasite have MPD of 0
names(nearctic.mam.mpd)<-nearctic_mamlist

## Create MPD scores for every host using the helminth tree
nearctic.mam.mpd.remove <- mpd(t(nearctic_df), nearctic_dis_para)
names(nearctic.mam.mpd.remove)<-nearctic_mamlist
nearctic.mam.mpd.remove<-nearctic.mam.mpd.remove[!is.na(nearctic.mam.mpd.remove)] ## hosts infected by only one parasite have MPD of 0


##calculate the mpd of parasites and hosts!
global_dis_para<-cophenetic.phylo(global_tree) #interspecific distance matrix of parasites
global_dis_mam<-cophenetic.phylo(GMPD_mammal_tree) #intespecific distance matrix of mammals from small tree


##make the pipeline to get the absense presence of each parasite (rows) that infect a given host (column)
global_mamlist<-as.vector(GMPD_mammal_tree$tip.label) #list of mammals in tip label order
global_para_list<-as.vector(global_tree$tip.label)
global_df <- data.frame(matrix(ncol=length(global_mamlist),nrow=length(global_para_list)), row.names = global_para_list)
colnames(global_df)<-global_mamlist

for (i in 1:nrow(global_df))
  global_df[i,] <- (colnames(global_df) %in% unique(subset(GMPDsmall, ParasiteCorrectedName==rownames(global_df)[i])$HostCorrectedName)) %>% as.numeric

write.csv(global_df,file = "global_df.csv", row.names = TRUE, col.names = TRUE)

## Create MPD scores for every parasite using the Phylacine tree
global.para.mpd <- mpd(global_df, global_dis_mam)
global.para.mpd[is.na(global.para.mpd)] <- 0 ## one host parasites have an mpd of 0
names(global.para.mpd)<-global_para_list

global.para.mpd.remove <- mpd(global_df, global_dis_mam)
names(global.para.mpd.remove)<-global_para_list
global.para.mpd.remove<-global.para.mpd.remove[!is.na(global.para.mpd.remove)] ## one host parasites are removed

## Create MPD scores for every host using the helminth tree
global.mam.mpd <- mpd(t(global_df), global_dis_para)
global.mam.mpd[is.na(global.mam.mpd)] <- 0 ## hosts infected by only one parasite have MPD of 0
names(global.mam.mpd)<-global_mamlist

## Create MPD scores for every host using the helminth tree
global.mam.mpd.remove <- mpd(t(global_df), global_dis_para)
names(global.mam.mpd.remove)<-global_mamlist
global.mam.mpd.remove<-global.mam.mpd.remove[!is.na(global.mam.mpd.remove)] ## hosts infected by only one parasite have MPD of 0

```




Nearctic helminth phylogeny, with single host taxa's MPD value = 0, branches colored by MPD value
```{r echo = FALSE}
contMap(nearctic_tree, nearctic.para.mpd, fsize= .3, lwd=1, leg.text="x", type="fan") #90 species
```





Test of phylogenetic signal in MPD using the Nearctic helminth phylogeny, with single host taxa's MPD value = 0
```{r echo = FALSE}
##test phylogenetic signal of mpd with single host species = 0
#fitContinuous(nearctic_tree, nearctic.para.mpd, model="lambda") 
#lambda = 0.384 ; aicc = 1017.13
#fitContinuous(nearctic_tree, nearctic.para.mpd, model="white")  
#                 aicc = 1023.17           



nearctic_mpd_evo<-c(1017.13,1023.17)
names(nearctic_mpd_evo)<-c("lambda","star")
nearctic_aic_mpd_signal <- tibble(aicc=nearctic_mpd_evo, model=names(nearctic_mpd_evo))
nearctic_aic_mpd<-nearctic_aic_mpd_signal %>% 
   mutate(deltaaicc=aicc-min(aicc), 
          pow=exp(-.5*deltaaicc),
          sumpow=sum(pow),
          waicc=pow/sumpow) %>%
   dplyr::select(c(2,1,3,6)) %>%
   arrange(desc(waicc))

nearctic_aic_mpd
```



\newpage
Nearctic helminth phylogney, with single host taxa removed, branches colored by MPD value
```{r echo = FALSE}
#view mpd with single host species removed on the tree
setdiff(nearctic_tree$tip.label, names(nearctic.para.mpd.remove))->bye4
remove_nearctic_tree<-drop.tip(nearctic_tree, nearctic_tree$tip.label[which(nearctic_tree$tip.label%in%bye4)])
#plot(remove_nearctic_tree, cex=0.3)

contMap(remove_nearctic_tree, nearctic.para.mpd.remove, fsize=.3, lwd=1) #70 species

```




Test of phylogenetic signal in MPD using the Nearctic helminth phylogeny,with single host taxa removed
```{r echo = FALSE}
##test phylogenetic signal of mpd with single host species removed
#fitContinuous(remove_nearctic_tree, nearctic.para.mpd.remove, model="lambda") 
#lambda = 0.926; aicc = 773.37
#fitContinuous(remove_nearctic_tree, nearctic.para.mpd.remove, model="white")  
#                aicc = 791.72


remove_nearctic_mpd_evo<-c(773.37,791.72)
names(remove_nearctic_mpd_evo)<-c("lambda","star")

remove_nearctic_aic_mpd_signal <- tibble(aicc=remove_nearctic_mpd_evo, model=names(remove_nearctic_mpd_evo))

removed_nearctic_aic_mpd<-remove_nearctic_aic_mpd_signal %>% 
  mutate(deltaaicc=aicc-min(aicc), 
         pow=exp(-.5*deltaaicc),
         sumpow=sum(pow),
         waicc=pow/sumpow) %>%
  dplyr::select(c(2,1,3,6)) %>%
  arrange(desc(waicc))

removed_nearctic_aic_mpd
```



\newpage
GMPD helminth phylogeny, with single host taxa's MPD value = 0, branches colored by MPD value
```{r echo = FALSE}
#view mpd with single host species = 0 on the tree
contMap(global_tree, global.para.mpd, fsize= .3,lwd=2, type="fan") #249 species
```





Test of phylogenetic signal in MPD using the GMPD helminth phylogeny, with single host taxa's MPD value = 0
```{r echo = FALSE}
##test phylogenetic signal of mpd with single host species = 0
#fitContinuous(global_tree, global.para.mpd, model="lambda") 
#lambda = 0.607 ; aicc = 2473.54
#fitContinuous(global_tree, global.para.mpd, model="white")  
#                 aicc = 2509.03           


global_mpd_evo<-c(2473.54,2509.03)
names(global_mpd_evo)<-c("lambda","star")
global_aic_mpd_signal <- tibble(aicc=global_mpd_evo, model=names(global_mpd_evo))
global_aic_mpd<-global_aic_mpd_signal %>% 
   mutate(deltaaicc=aicc-min(aicc), 
          pow=exp(-.5*deltaaicc),
          sumpow=sum(pow),
          waicc=pow/sumpow) %>%
   dplyr::select(c(2,1,3,6)) %>%
   arrange(desc(waicc))

global_aic_mpd
```



\newpage
GMPD helminth phylogeny, with single host taxa removed, branches colored by MPD value
```{r echo = FALSE}
#view mpd with single host species removed on the tree
setdiff(global_tree$tip.label, names(global.para.mpd.remove))->bye5
remove_global_tree<-drop.tip(global_tree, global_tree$tip.label[which(global_tree$tip.label%in%bye5)])
#plot(remove_global_tree, cex=0.3)

contMap(remove_global_tree, global.para.mpd.remove, fsize=.3, lwd=1, type="fan") #159 species
```





Test of phylogenetic signal in MPD using the GMPD helminth phylogeny, with single host taxa removed
```{r echo = FALSE}
##test phylogenetic signal of mpd with single host species removed
#fitContinuous(remove_global_tree, global.para.mpd.remove, model="lambda") 
#lambda = 0.895; aicc = 1547.52
#fitContinuous(remove_global_tree, global.para.mpd.remove, model="white")  
#                aicc = 1606.16


remove_global_mpd_evo<-c(1547.52,1606.16)
names(remove_global_mpd_evo)<-c("lambda","star")

remove_global_aic_mpd_signal <- tibble(aicc=remove_global_mpd_evo, model=names(remove_global_mpd_evo))

removed_global_aic_mpd<-remove_global_aic_mpd_signal %>% 
  mutate(deltaaicc=aicc-min(aicc), 
         pow=exp(-.5*deltaaicc),
         sumpow=sum(pow),
         waicc=pow/sumpow) %>%
  dplyr::select(c(2,1,3,6)) %>%
  arrange(desc(waicc))

removed_global_aic_mpd
```


\newpage
Test of phylogenetic signal in taxonomic breadth using the Nearctic helminth phylogeny, with single host taxa included
```{r echo = FALSE}
##same order as tree tip labels
nearctic_freq<- nearctic_abund_data[match(nearctic_tree$tip.label, nearctic_abund_data$Parasite),]

nearctic.no.taxa<-nearctic_freq$Freq #includes single host species
names(nearctic.no.taxa)<-nearctic_freq$Parasite


remove.nearctic.no.taxa<-(nearctic.no.taxa[!nearctic.no.taxa == 1]) # removes single host species

##test phylogenetic signal for data with single host species
#fitContinuous(nearctic_tree, nearctic.no.taxa, model="lambda") 
#lambda=0.094,  aicc= 471.32
#fitContinuous(nearctic_tree, nearctic.no.taxa, model="white")                 
#               aicc= 469.62


nearctic_no_taxa_evo<-c(471.32, 469.62)
names(nearctic_no_taxa_evo)<-c("lambda", "star")

nearctic_aic_taxa <- tibble(aicc=nearctic_no_taxa_evo, model=names(nearctic_no_taxa_evo))

nearctic_aic_taxa<-nearctic_aic_taxa %>% 
  mutate(deltaaicc=aicc-min(aicc), 
         pow=exp(-.5*deltaaicc),
         sumpow=sum(pow),
         waicc=pow/sumpow) %>%
  dplyr::select(c(2,1,3,6)) %>%
  arrange(desc(waicc))

nearctic_aic_taxa

```





Test of phylogenetic signal in taxonomic breadth using the Nearctic helminth phylogeny, with single host taxa removed
```{r echo = FALSE}
##test phylogenetic signal for data without single host species
#fitContinuous(remove_nearctic_tree, remove.nearctic.no.taxa, model="lambda") 
#lambda=0.091,  aicc= 363.06
#fitContinuous(remove_nearctic_tree, remove.nearctic.no.taxa, model="white")              
#               aicc= 361.07

remove_nearctic_no_taxa_evo<-c(363.06, 361.07)
names(remove_nearctic_no_taxa_evo)<-c("lambda", "star")

remove_nearctic_aic_taxa <- tibble(aicc=remove_nearctic_no_taxa_evo, model=names(remove_nearctic_no_taxa_evo))

remove_nearctic_aic_taxa<-remove_nearctic_aic_taxa %>% 
  mutate(deltaaicc=aicc-min(aicc), 
         pow=exp(-.5*deltaaicc),
         sumpow=sum(pow),
         waicc=pow/sumpow) %>%
  dplyr::select(c(2,1,3,6)) %>%
  arrange(desc(waicc))

remove_nearctic_aic_taxa
```





Test of phylogenetic signal in taxonomic breadth using the GMPD helminth phylogeny, with single host taxa included
```{r echo = FALSE}
load("get_nriABUND.Rda") ## nri

no_taxa<-as.data.frame(nri$ntaxa)
no_taxa$parasite<-rownames(nri)

global_freq<-no_taxa[no_taxa$parasite %in% global_tree$tip.label, ] #197 species

#subset global tree 
setdiff(global_tree$tip.label, no_taxa$parasite)->bye6
ntaxa_global_tree<-drop.tip(global_tree, global_tree$tip.label[which(global_tree$tip.label%in%bye6)])
#check
#plot(ntaxa_global_tree, cex=0.3)

##same order as tree tip labels
g_freq<- global_freq[match(ntaxa_global_tree$tip.label, global_freq$parasite),]

#make the vector a named number
global.no.taxa<-g_freq$`nri$ntaxa` #includes single host species
names(global.no.taxa)<-g_freq$parasite

#check
#contMap(ntaxa_global_tree, global.no.taxa, fsize=.3, type = "fan")


#remove single host species
remove.global.no.taxa<-(global.no.taxa[!global.no.taxa == 1]) #134 species

#subset a tree to match the removed dataset
setdiff(ntaxa_global_tree$tip.label, names(remove.global.no.taxa))->bye7
remove_ntaxa_global_tree<-drop.tip(ntaxa_global_tree, ntaxa_global_tree$tip.label[which(ntaxa_global_tree$tip.label%in%bye7)])

#check
#contMap(remove_ntaxa_global_tree, remove.global.no.taxa, fsize=.3, type ="fan")



##test phylogenetic signal for data with single host species
#fitContinuous(ntaxa_global_tree, global.no.taxa, model="lambda") 
#lambda=0.000,  aicc= 1159.04
#fitContinuous(ntaxa_global_tree, global.no.taxa, model="white")                 
#               aicc= 1156.97


global_no_taxa_evo<-c(1159.04, 1156.97)
names(global_no_taxa_evo)<-c("lambda", "star")

global_aic_taxa <- tibble(aicc=global_no_taxa_evo, model=names(global_no_taxa_evo))

global_aic_taxa<-global_aic_taxa %>% 
  mutate(deltaaicc=aicc-min(aicc), 
         pow=exp(-.5*deltaaicc),
         sumpow=sum(pow),
         waicc=pow/sumpow) %>%
  dplyr::select(c(2,1,3,6)) %>%
  arrange(desc(waicc))

global_aic_taxa
```





Test of phylogenetic signal in taxonomic breadth using the GMPD helminth phylogeny, with single host taxa removed
```{r echo = FALSE}
##test phylogenetic signal for data without single host species
#fitContinuous(remove_ntaxa_global_tree, remove.global.no.taxa, model="lambda") 
#lambda=0.000,  aicc = 802.86
#fitContinuous(remove_ntaxa_global_tree, remove.global.no.taxa, model="white")             
#               aicc = 800.77

remove_global_no_taxa_evo<-c(802.86, 800.77)
names(remove_global_no_taxa_evo)<-c("lambda", "star")

remove_global_aic_taxa <- tibble(aicc=remove_global_no_taxa_evo, model=names(remove_global_no_taxa_evo))

remove_global_aic_taxa<-remove_global_aic_taxa %>% 
  mutate(deltaaicc=aicc-min(aicc), 
         pow=exp(-.5*deltaaicc),
         sumpow=sum(pow),
         waicc=pow/sumpow) %>%
  dplyr::select(c(2,1,3,6)) %>%
  arrange(desc(waicc))

remove_global_aic_taxa
```


```{r include = FALSE}
###NEARCTIC DATASET####

#match GMPD data to the nearctic tree (to get Phyla classifications)
g<-GMPDsmall[match(nearctic_tree$tip.label, GMPDsmall$ParasiteCorrectedName),]

#make Phyla dataframe (names, regimes, trait)
nearctic_phyla_data<-data.frame(nearctic_tree$tip.label)
nearctic_phyla_data$regime<-g$ParPhylum
nearctic_phyla_data$trait<-nearctic.para.mpd

## remake to be able to use ouch rather than OUwie
nearctic_ouchtree <- ape2ouch(nearctic_tree) ## convert to ouchtree object
## get the mpd data organized properly for ouch
nearctic_ouchdata <- rep(NA, length=nearctic_ouchtree@nnodes)
for (i in 1:length(nearctic_phyla_data$nearctic_tree.tip.label))
  nearctic_ouchdata[which(nearctic_ouchtree@nodelabels==nearctic_phyla_data$nearctic_tree.tip.label[i])] <- nearctic_phyla_data$trait[i]
nearctic_ouchdata <- as.data.frame(nearctic_ouchdata)

## Find the node that corresponds to the most recent common ancestor of all Platyhelminthes
## which nodes in the ouchtree correspond to the Plats
nearctic_ouchtree@nodelabels%in%nearctic_phyla_data$nearctic_tree.tip.label[which(nearctic_phyla_data$regime=="Platyhelminthes")] %>% which -> platys
## get their lineages and find the nodes that are in common across all lineages
nearctic_ouchtree@lineages[platys] -> platy_lineages
platy_lineages[[1]][sapply(platy_lineages[[1]], function(n) lapply(platy_lineages, function(l) n%in%l) %>% unlist %>% all) %>% which %>% min] ## 1

## Find the node that corresponds to the most recent common ancestor of all Acanthocephalans
## which nodes in the ouchtree correspond to the Acanths
nearctic_ouchtree@nodelabels%in%nearctic_phyla_data$nearctic_tree.tip.label[which(nearctic_phyla_data$regime=="Acanthocephala")] %>% which -> acanths
## get their lineages and find the nodes that are in common across all lineages
nearctic_ouchtree@lineages[acanths] -> acanth_lineages
acanth_lineages[[1]][sapply(acanth_lineages[[1]], function(n) lapply(acanth_lineages, function(l) n%in%l) %>% unlist %>% all) %>% which %>% min] ## 86

## Find the node that corresponds to the most recent common ancestor of all Nematodes
## which nodes in the ouchtree correspond to the Nematodes
nearctic_ouchtree@nodelabels%in%nearctic_phyla_data$nearctic_tree.tip.label[which(nearctic_phyla_data$regime=="Nematoda")] %>% which -> nematodes
## get their lineages and find the nodes that are in common across all lineages
nearctic_ouchtree@lineages[nematodes] -> nematode_lineages
nematode_lineages[[1]][sapply(nematode_lineages[[1]], function(n) lapply(nematode_lineages, function(l) n%in%l) %>% unlist %>% all) %>% which %>% min]

## paint the taxonomy-based regimes onto the tree
paint(nearctic_ouchtree, subtree=c("1"="Platys","86"="Acanths","81"="Nematodes"), branch=c("1"="Platys")) -> nearctic_oum_regimes
## paint a single global regime onto the tree
paint(nearctic_ouchtree, subtree=c("1"="global"), branch=c("1"="global")) -> nearctic_ou1_regimes

## Plot to make sure that you end up with a reasonable painting
plot(nearctic_ouchtree, regimes=nearctic_oum_regimes)

## Now do the fitting using ouch
## OUM: AICc = 1008.5, alpha = 21.2, sigma.sq = 167044, Acanths = 134, Nematodes = 41.8, Platys = 95.2
nearctic_oum_ouch <- hansen(data=nearctic_ouchdata, tree=nearctic_ouchtree, regimes=nearctic_oum_regimes, sqrt.alpha=100, sigma=100)

## OU1: AICc = 1017.7, alpha = 6.35, sigma.sq = 63148, global = 71.7
nearctic_ou1_ouch <- hansen(data=nearctic_ouchdata, tree=nearctic_ouchtree, regimes=nearctic_ou1_regimes, sqrt.alpha=100, sigma=100)

## BM: AICc = 1039.0, sigma.sq = 25897
nearctic_bm_ouch <- brown(data=nearctic_ouchdata, tree=nearctic_ouchtree)

#######################################


## Nearctic dataset with single host species MPD removed

#match GMPD data to the remove_nearctic tree (to get Phyla classifications)
h<-GMPDsmall[match(remove_nearctic_tree$tip.label, GMPDsmall$ParasiteCorrectedName),]

#make Phyla dataframe (names, regimes, trait)
remove_nearctic_phyla_data<-data.frame(remove_nearctic_tree$tip.label)
remove_nearctic_phyla_data$regime<-h$ParPhylum
remove_nearctic_phyla_data$trait<-nearctic.para.mpd.remove

## remake to be able to use ouch rather than OUwie
remove_nearctic_ouchtree <- ape2ouch(remove_nearctic_tree) ## convert to ouchtree object
## get the mpd data organized properly for ouch
remove_nearctic_ouchdata <- rep(NA, length=remove_nearctic_ouchtree@nnodes)
for (i in 1:length(remove_nearctic_phyla_data$remove_nearctic_tree.tip.label))
  remove_nearctic_ouchdata[which(remove_nearctic_ouchtree@nodelabels==remove_nearctic_phyla_data$remove_nearctic_tree.tip.label[i])] <- remove_nearctic_phyla_data$trait[i]
remove_nearctic_ouchdata <- as.data.frame(remove_nearctic_ouchdata)

## You can use the code above, replacing nearctic with remove_nearctic, to find the last common ancestor for Platyhelminthes, Acanthocephalans, and Nematodes. I drop those blocks of code here to reduce the amount of code.
## paint the regimes onto the tree
paint(remove_nearctic_ouchtree, subtree=c("1"="Platys","67"="Acanths","64"="Nematodes"), branch=c("1"="Platys")) -> remove_nearctic_oum_regimes
paint(remove_nearctic_ouchtree, subtree=c("1"="global"), branch=c("1"="global")) -> remove_nearctic_ou1_regimes
## Plot it to make sure it looks good
plot(remove_nearctic_ouchtree, regimes=remove_nearctic_oum_regimes)

## Now do the fitting using ouch

## OUM: AICc = 762.9, alpha = 458.6, sigma.sq = 2628906, Acanths = 201, Nematodes = 52.8, Platys = 119
remove_nearctic_oum_ouch <- hansen(data=remove_nearctic_ouchdata, tree=remove_nearctic_ouchtree, regimes=remove_nearctic_oum_regimes, sqrt.alpha=100, sigma=100)

## OU1: AICc = 778.5, alpha = 2.63, sigma.sq = 26815.8, global = 99.04
remove_nearctic_ou1_ouch <- hansen(data=remove_nearctic_ouchdata, tree=remove_nearctic_ouchtree, regimes=remove_nearctic_ou1_regimes, sqrt.alpha=100, sigma=100)

## BM: AICc = 782.6, sigma.sq = 15840
remove_nearctic_bm_ouch <- brown(data=remove_nearctic_ouchdata, tree=remove_nearctic_ouchtree)

#######################################

###GLOBAL DATASET####

#match GMPD data to the global tree
k<-GMPDsmall[match(global_tree$tip.label, GMPDsmall$ParasiteCorrectedName),]

#make Phyla dataframe (names, regimes, trait)
global_phyla_data<-data.frame(global_tree$tip.label)
global_phyla_data$regime<-k$ParPhylum
global_phyla_data$trait<-global.para.mpd

## Painting taxonomy-based regimes onto this tree
global_ouchtree <- ape2ouch(global_tree) ## convert to ouchtree object
## get the mpd data organized properly for ouch
global_ouchdata <- rep(NA, length=global_ouchtree@nnodes)
for (i in 1:length(global_phyla_data$global_tree.tip.label))
  global_ouchdata[which(global_ouchtree@nodelabels==global_phyla_data$global_tree.tip.label[i])] <- global_phyla_data$trait[i]
global_ouchdata <- as.data.frame(global_ouchdata)

## You can use the code above, replacing nearctic with global, to find the last common ancestor for Platyhelminthes, Acanthocephalans, and Nematodes. I drop those blocks of code here to reduce the amount of code.

## paint the regimes onto the tree
paint(global_ouchtree, subtree=c("1"="Platys","227"="Nematodes","243"="Acanths"), branch=c("1"="Platys")) -> global_oum_regimes
## global regime
paint(global_ouchtree, subtree=c("1"="global"), branch=c("1"="global")) -> global_ou1_regimes
## Plot it to make sure it looks good
plot(global_ouchtree, regimes=global_oum_regimes)


## Now do the fitting using ouch
## OUM: AICc = 2480.3, alpha = 16.8, sigma.sq = 42480.8, Acanths = 24.8, Nematodes = 21.7, Platys = 47.6
global_oum_ouch <- hansen(data=global_ouchdata, tree=global_ouchtree, regimes=global_oum_regimes, sqrt.alpha=100, sigma=100)

## OU1: AICc = 2490.3, alpha = 9.14, sigma.sq = 26943.6, global = 33.3 
global_ou1_ouch <- hansen(data=global_ouchdata, tree=global_ouchtree, regimes=global_ou1_regimes, sqrt.alpha=100, sigma=100)

## BM: AICc = 2571.44, sigma.sq = 12334.6
global_bm_ouch <- brown(data=global_ouchdata, tree=global_ouchtree)

#######################################

##with single host species MPD removed

#match GMPD data to the remove_global tree
l<-GMPDsmall[match(remove_global_tree$tip.label, GMPDsmall$ParasiteCorrectedName),]

#make Phyla dataframe for OUwie (names, regimes, trait)
remove_global_phyla_data<-data.frame(remove_global_tree$tip.label)
remove_global_phyla_data$regime<-l$ParPhylum
remove_global_phyla_data$trait<-global.para.mpd.remove

## Prepare for ouch
remove_global_ouchtree <- ape2ouch(remove_global_tree) ## convert to ouchtree object
## get the mpd data organized properly for ouch
remove_global_ouchdata <- rep(NA, length=remove_global_ouchtree@nnodes)
for (i in 1:length(remove_global_phyla_data$remove_global_tree.tip.label))
  remove_global_ouchdata[which(remove_global_ouchtree@nodelabels==remove_global_phyla_data$remove_global_tree.tip.label[i])] <- remove_global_phyla_data$trait[i]
remove_global_ouchdata <- as.data.frame(remove_global_ouchdata)

## You can use the code above, replacing nearctic with remove_global, to find the last common ancestor for Platyhelminthes, Acanthocephalans, and Nematodes. I drop those blocks of code here to reduce the amount of code.


## paint the regimes onto the tree
paint(remove_global_ouchtree, subtree=c("1"="Platys","150"="Nematodes","155"="Acanths"), branch=c("1"="Platys")) -> remove_global_oum_regimes
## global regime
paint(remove_global_ouchtree, subtree=c("1"="global"), branch=c("1"="global")) -> remove_global_ou1_regimes

## Plot it to make sure it looks good
plot(remove_global_ouchtree, regimes=remove_global_oum_regimes)
###NEARCTIC DATASET####

#match GMPD data to the nearctic tree (to get Phyla classifications)
g<-GMPDsmall[match(nearctic_tree$tip.label, GMPDsmall$ParasiteCorrectedName),]

#make Phyla dataframe (names, regimes, trait)
nearctic_phyla_data<-data.frame(nearctic_tree$tip.label)
nearctic_phyla_data$regime<-g$ParPhylum
nearctic_phyla_data$trait<-nearctic.para.mpd

## remake to be able to use ouch rather than OUwie
nearctic_ouchtree <- ape2ouch(nearctic_tree) ## convert to ouchtree object
## get the mpd data organized properly for ouch
nearctic_ouchdata <- rep(NA, length=nearctic_ouchtree@nnodes)
for (i in 1:length(nearctic_phyla_data$nearctic_tree.tip.label))
  nearctic_ouchdata[which(nearctic_ouchtree@nodelabels==nearctic_phyla_data$nearctic_tree.tip.label[i])] <- nearctic_phyla_data$trait[i]
nearctic_ouchdata <- as.data.frame(nearctic_ouchdata)

## Find the node that corresponds to the most recent common ancestor of all Platyhelminthes
## which nodes in the ouchtree correspond to the Plats
nearctic_ouchtree@nodelabels%in%nearctic_phyla_data$nearctic_tree.tip.label[which(nearctic_phyla_data$regime=="Platyhelminthes")] %>% which -> platys
## get their lineages and find the nodes that are in common across all lineages
nearctic_ouchtree@lineages[platys] -> platy_lineages
platy_lineages[[1]][sapply(platy_lineages[[1]], function(n) lapply(platy_lineages, function(l) n%in%l) %>% unlist %>% all) %>% which %>% min] ## 1

## Find the node that corresponds to the most recent common ancestor of all Acanthocephalans
## which nodes in the ouchtree correspond to the Acanths
nearctic_ouchtree@nodelabels%in%nearctic_phyla_data$nearctic_tree.tip.label[which(nearctic_phyla_data$regime=="Acanthocephala")] %>% which -> acanths
## get their lineages and find the nodes that are in common across all lineages
nearctic_ouchtree@lineages[acanths] -> acanth_lineages
acanth_lineages[[1]][sapply(acanth_lineages[[1]], function(n) lapply(acanth_lineages, function(l) n%in%l) %>% unlist %>% all) %>% which %>% min] ## 86

## Find the node that corresponds to the most recent common ancestor of all Nematodes
## which nodes in the ouchtree correspond to the Nematodes
nearctic_ouchtree@nodelabels%in%nearctic_phyla_data$nearctic_tree.tip.label[which(nearctic_phyla_data$regime=="Nematoda")] %>% which -> nematodes
## get their lineages and find the nodes that are in common across all lineages
nearctic_ouchtree@lineages[nematodes] -> nematode_lineages
nematode_lineages[[1]][sapply(nematode_lineages[[1]], function(n) lapply(nematode_lineages, function(l) n%in%l) %>% unlist %>% all) %>% which %>% min]

## paint the taxonomy-based regimes onto the tree
paint(nearctic_ouchtree, subtree=c("1"="Platys","86"="Acanths","81"="Nematodes"), branch=c("1"="Platys")) -> nearctic_oum_regimes
## paint a single global regime onto the tree
paint(nearctic_ouchtree, subtree=c("1"="global"), branch=c("1"="global")) -> nearctic_ou1_regimes

## Plot to make sure that you end up with a reasonable painting
plot(nearctic_ouchtree, regimes=nearctic_oum_regimes)

## Now do the fitting using ouch
## OUM: AICc = 1008.5, alpha = 21.2, sigma.sq = 167044, Acanths = 134, Nematodes = 41.8, Platys = 95.2
nearctic_oum_ouch <- hansen(data=nearctic_ouchdata, tree=nearctic_ouchtree, regimes=nearctic_oum_regimes, sqrt.alpha=100, sigma=100)

## OU1: AICc = 1017.7, alpha = 6.35, sigma.sq = 63148, global = 71.7
nearctic_ou1_ouch <- hansen(data=nearctic_ouchdata, tree=nearctic_ouchtree, regimes=nearctic_ou1_regimes, sqrt.alpha=100, sigma=100)

## BM: AICc = 1039.0, sigma.sq = 25897
nearctic_bm_ouch <- brown(data=nearctic_ouchdata, tree=nearctic_ouchtree)

#######################################


## Nearctic dataset with single host species MPD removed

#match GMPD data to the remove_nearctic tree (to get Phyla classifications)
h<-GMPDsmall[match(remove_nearctic_tree$tip.label, GMPDsmall$ParasiteCorrectedName),]

#make Phyla dataframe (names, regimes, trait)
remove_nearctic_phyla_data<-data.frame(remove_nearctic_tree$tip.label)
remove_nearctic_phyla_data$regime<-h$ParPhylum
remove_nearctic_phyla_data$trait<-nearctic.para.mpd.remove

## remake to be able to use ouch rather than OUwie
remove_nearctic_ouchtree <- ape2ouch(remove_nearctic_tree) ## convert to ouchtree object
## get the mpd data organized properly for ouch
remove_nearctic_ouchdata <- rep(NA, length=remove_nearctic_ouchtree@nnodes)
for (i in 1:length(remove_nearctic_phyla_data$remove_nearctic_tree.tip.label))
  remove_nearctic_ouchdata[which(remove_nearctic_ouchtree@nodelabels==remove_nearctic_phyla_data$remove_nearctic_tree.tip.label[i])] <- remove_nearctic_phyla_data$trait[i]
remove_nearctic_ouchdata <- as.data.frame(remove_nearctic_ouchdata)

## You can use the code above, replacing nearctic with remove_nearctic, to find the last common ancestor for Platyhelminthes, Acanthocephalans, and Nematodes. I drop those blocks of code here to reduce the amount of code.
## paint the regimes onto the tree
paint(remove_nearctic_ouchtree, subtree=c("1"="Platys","67"="Acanths","64"="Nematodes"), branch=c("1"="Platys")) -> remove_nearctic_oum_regimes
paint(remove_nearctic_ouchtree, subtree=c("1"="global"), branch=c("1"="global")) -> remove_nearctic_ou1_regimes
## Plot it to make sure it looks good
plot(remove_nearctic_ouchtree, regimes=remove_nearctic_oum_regimes)

## Now do the fitting using ouch

## OUM: AICc = 762.9, alpha = 458.6, sigma.sq = 2628906, Acanths = 201, Nematodes = 52.8, Platys = 119
remove_nearctic_oum_ouch <- hansen(data=remove_nearctic_ouchdata, tree=remove_nearctic_ouchtree, regimes=remove_nearctic_oum_regimes, sqrt.alpha=100, sigma=100)

## OU1: AICc = 778.5, alpha = 2.63, sigma.sq = 26815.8, global = 99.04
remove_nearctic_ou1_ouch <- hansen(data=remove_nearctic_ouchdata, tree=remove_nearctic_ouchtree, regimes=remove_nearctic_ou1_regimes, sqrt.alpha=100, sigma=100)

## BM: AICc = 782.6, sigma.sq = 15840
remove_nearctic_bm_ouch <- brown(data=remove_nearctic_ouchdata, tree=remove_nearctic_ouchtree)

#######################################

###GLOBAL DATASET####

#match GMPD data to the global tree
k<-GMPDsmall[match(global_tree$tip.label, GMPDsmall$ParasiteCorrectedName),]

#make Phyla dataframe (names, regimes, trait)
global_phyla_data<-data.frame(global_tree$tip.label)
global_phyla_data$regime<-k$ParPhylum
global_phyla_data$trait<-global.para.mpd

## Painting taxonomy-based regimes onto this tree
global_ouchtree <- ape2ouch(global_tree) ## convert to ouchtree object
## get the mpd data organized properly for ouch
global_ouchdata <- rep(NA, length=global_ouchtree@nnodes)
for (i in 1:length(global_phyla_data$global_tree.tip.label))
  global_ouchdata[which(global_ouchtree@nodelabels==global_phyla_data$global_tree.tip.label[i])] <- global_phyla_data$trait[i]
global_ouchdata <- as.data.frame(global_ouchdata)

## You can use the code above, replacing nearctic with global, to find the last common ancestor for Platyhelminthes, Acanthocephalans, and Nematodes. I drop those blocks of code here to reduce the amount of code.

## paint the regimes onto the tree
paint(global_ouchtree, subtree=c("1"="Platys","227"="Nematodes","243"="Acanths"), branch=c("1"="Platys")) -> global_oum_regimes
## global regime
paint(global_ouchtree, subtree=c("1"="global"), branch=c("1"="global")) -> global_ou1_regimes
## Plot it to make sure it looks good
plot(global_ouchtree, regimes=global_oum_regimes)

## Now do the fitting using ouch
## OUM: AICc = 2480.3, alpha = 16.8, sigma.sq = 42480.8, Acanths = 24.8, Nematodes = 21.7, Platys = 47.6
global_oum_ouch <- hansen(data=global_ouchdata, tree=global_ouchtree, regimes=global_oum_regimes, sqrt.alpha=100, sigma=100)

## OU1: AICc = 2490.3, alpha = 9.14, sigma.sq = 26943.6, global = 33.3 
global_ou1_ouch <- hansen(data=global_ouchdata, tree=global_ouchtree, regimes=global_ou1_regimes, sqrt.alpha=100, sigma=100)

## BM: AICc = 2571.44, sigma.sq = 12334.6
global_bm_ouch <- brown(data=global_ouchdata, tree=global_ouchtree)

#######################################

##with single host species MPD removed

#match GMPD data to the remove_global tree
l<-GMPDsmall[match(remove_global_tree$tip.label, GMPDsmall$ParasiteCorrectedName),]

#make Phyla dataframe for OUwie (names, regimes, trait)
remove_global_phyla_data<-data.frame(remove_global_tree$tip.label)
remove_global_phyla_data$regime<-l$ParPhylum
remove_global_phyla_data$trait<-global.para.mpd.remove

## Prepare for ouch
remove_global_ouchtree <- ape2ouch(remove_global_tree) ## convert to ouchtree object
## get the mpd data organized properly for ouch
remove_global_ouchdata <- rep(NA, length=remove_global_ouchtree@nnodes)
for (i in 1:length(remove_global_phyla_data$remove_global_tree.tip.label))
  remove_global_ouchdata[which(remove_global_ouchtree@nodelabels==remove_global_phyla_data$remove_global_tree.tip.label[i])] <- remove_global_phyla_data$trait[i]
remove_global_ouchdata <- as.data.frame(remove_global_ouchdata)

## You can use the code above, replacing nearctic with remove_global, to find the last common ancestor for Platyhelminthes, Acanthocephalans, and Nematodes. I drop those blocks of code here to reduce the amount of code.


## paint the regimes onto the tree
paint(remove_global_ouchtree, subtree=c("1"="Platys","150"="Nematodes","155"="Acanths"), branch=c("1"="Platys")) -> remove_global_oum_regimes
## global regime
paint(remove_global_ouchtree, subtree=c("1"="global"), branch=c("1"="global")) -> remove_global_ou1_regimes

## Plot it to make sure it looks good
plot(remove_global_ouchtree, regimes=remove_global_oum_regimes)

## Now do the fitting using ouch
## OUM: AICc = 1539.6, alpha = 6.49, sigma.sq = 13769.8, Acanths = 67.7, Nematodes = 37.5, Platys = 68.2
remove_global_oum_ouch <- hansen(data=remove_global_ouchdata, tree=remove_global_ouchtree, regimes=remove_global_oum_regimes, sqrt.alpha=100, sigma=100)

## OU1: AICc = 1543.6, alpha = 4.55, sigma.sq = 11222.8, global = 54.8 
remove_global_ou1_ouch <- hansen(data=remove_global_ouchdata, tree=remove_global_ouchtree, regimes=remove_global_ou1_regimes, sqrt.alpha=100, sigma=100)

## BM: AICc = 1563.4, sigma.sq = 6267.8
remove_global_bm_ouch <- brown(data=remove_global_ouchdata, tree=remove_global_ouchtree)

## Load all of the bootstrap data
fit_oum_model_to_oum_data_nearctic <- readRDS(file="bootstrap_results/fit_oum_model_to_oum_data_nearctic.RDS")
fit_oum_model_to_ou1_data_nearctic <- readRDS(file="bootstrap_results/fit_oum_model_to_ou1_data_nearctic.RDS")
fit_oum_model_to_bm_data_nearctic <- readRDS(file="bootstrap_results/fit_oum_model_to_bm_data_nearctic.RDS")
fit_ou1_model_to_oum_data_nearctic <- readRDS(file="bootstrap_results/fit_ou1_model_to_oum_data_nearctic.RDS")
fit_ou1_model_to_ou1_data_nearctic <- readRDS(file="bootstrap_results/fit_ou1_model_to_ou1_data_nearctic.RDS")
fit_bm_model_to_oum_data_nearctic <- readRDS(file="bootstrap_results/fit_bm_model_to_oum_data_nearctic.RDS")
fit_bm_model_to_bm_data_nearctic <- readRDS(file="bootstrap_results/fit_bm_model_to_bm_data_nearctic.RDS")

fit_oum_model_to_oum_data_remove_nearctic <- readRDS(file="bootstrap_results/fit_oum_model_to_oum_data_remove_nearctic.RDS")
fit_oum_model_to_ou1_data_remove_nearctic <- readRDS(file="bootstrap_results/fit_oum_model_to_ou1_data_remove_nearctic.RDS")
fit_oum_model_to_bm_data_remove_nearctic <- readRDS(file="bootstrap_results/fit_oum_model_to_bm_data_remove_nearctic.RDS")
fit_ou1_model_to_oum_data_remove_nearctic <- readRDS(file="bootstrap_results/fit_ou1_model_to_oum_data_remove_nearctic.RDS")
fit_ou1_model_to_ou1_data_remove_nearctic <- readRDS(file="bootstrap_results/fit_ou1_model_to_ou1_data_remove_nearctic.RDS")
fit_bm_model_to_oum_data_remove_nearctic <- readRDS(file="bootstrap_results/fit_bm_model_to_oum_data_remove_nearctic.RDS")
fit_bm_model_to_bm_data_remove_nearctic <- readRDS(file="bootstrap_results/fit_bm_model_to_bm_data_remove_nearctic.RDS")

fit_oum_model_to_oum_data_global <- readRDS(file="bootstrap_results/fit_oum_model_to_oum_data_global.RDS")
fit_oum_model_to_ou1_data_global <- readRDS(file="bootstrap_results/fit_oum_model_to_ou1_data_global.RDS")
fit_oum_model_to_bm_data_global <- readRDS(file="bootstrap_results/fit_oum_model_to_bm_data_global.RDS")
fit_ou1_model_to_oum_data_global <- readRDS(file="bootstrap_results/fit_ou1_model_to_oum_data_global.RDS")
fit_ou1_model_to_ou1_data_global <- readRDS(file="bootstrap_results/fit_ou1_model_to_ou1_data_global.RDS")
fit_bm_model_to_oum_data_global <- readRDS(file="bootstrap_results/fit_bm_model_to_oum_data_global.RDS")
fit_bm_model_to_bm_data_global <- readRDS(file="bootstrap_results/fit_bm_model_to_bm_data_global.RDS")

fit_oum_model_to_oum_data_remove_global <- readRDS(file="bootstrap_results/fit_oum_model_to_oum_data_remove_global.RDS")
fit_oum_model_to_ou1_data_remove_global <- readRDS(file="bootstrap_results/fit_oum_model_to_ou1_data_remove_global.RDS")
fit_oum_model_to_bm_data_remove_global <- readRDS(file="bootstrap_results/fit_oum_model_to_bm_data_remove_global.RDS")
fit_ou1_model_to_oum_data_remove_global <- readRDS(file="bootstrap_results/fit_ou1_model_to_oum_data_remove_global.RDS")
fit_ou1_model_to_ou1_data_remove_global <- readRDS(file="bootstrap_results/fit_ou1_model_to_ou1_data_remove_global.RDS")
fit_bm_model_to_oum_data_remove_global <- readRDS(file="bootstrap_results/fit_bm_model_to_oum_data_remove_global.RDS")
fit_bm_model_to_bm_data_remove_global <- readRDS(file="bootstrap_results/fit_bm_model_to_bm_data_remove_global.RDS")

## If we treat BM as the null model, then we can compute the likelihood differences observed when BM is true to the observed likelihood difference in the real data to determine a p-value
## 1. Compute the likelihood difference between the BM model fit to the BM data and the OU model fit to the BM data for all simulated datasets
## 2. Compute the likelihood difference between the BM model fit to the real data and the OU model fit to the real data
## 3. The number of likelihood differences computed in step #1 that are larger than the observed likelihood difference divided by the total number of simulated datasets gives an estimate of the p-value of the test (since the p-value is the probability of observing a dataset as extreme as the real one, assuming that the null model [BM] is true)
(((fit_bm_model_to_bm_data_global$logLik-fit_oum_model_to_bm_data_global$logLik) < (global_bm_ouch@loglik - global_oum_ouch@loglik)) %>% sum)/nrow(fit_bm_model_to_bm_data_global) -> p_value_bm_vs_oum_global #0 

(((fit_bm_model_to_bm_data_remove_global$logLik-fit_oum_model_to_bm_data_remove_global$logLik) < (remove_global_bm_ouch@loglik - remove_global_oum_ouch@loglik)) %>% sum)/nrow(fit_bm_model_to_bm_data_remove_global) -> p_value_bm_vs_oum_remove_global #0

(((fit_bm_model_to_bm_data_nearctic$logLik-fit_oum_model_to_bm_data_nearctic$logLik) < (nearctic_bm_ouch@loglik - nearctic_oum_ouch@loglik)) %>% sum)/nrow(fit_bm_model_to_bm_data_nearctic) -> p_value_bm_vs_oum_nearctic #0

(((fit_bm_model_to_bm_data_remove_nearctic$logLik-fit_oum_model_to_bm_data_remove_nearctic$logLik) < (remove_nearctic_bm_ouch@loglik - remove_nearctic_oum_ouch@loglik)) %>% sum)/nrow(fit_bm_model_to_bm_data_remove_nearctic) -> p_value_bm_vs_oum_remove_nearctic #0.002

## If we treat OU1 as the null model, then we can compute the likelihood differences observed when OU1 is true to the observed likelihood difference in the real data to determine a p-value
## 1. Compute the likelihood difference between the OU1 model fit to the OU1 data and the OUM model fit to the OU1 data for all simulated datasets
## 2. Compute the likelihood difference between the OU1 model fit to the real data and the OUM model fit to the real data
## 3. The number of likelihood differences computed in step #1 that are larger than the observed likelihood difference divided by the total number of simulated datasets gives an estimate of the p-value of the test (since the p-value is the probability of observing a dataset as extreme as the real one, assuming that the null model [BM] is true)
(((fit_ou1_model_to_ou1_data_global$logLik-fit_oum_model_to_ou1_data_global$logLik) < (global_ou1_ouch@loglik - global_oum_ouch@loglik)) %>% sum)/nrow(fit_ou1_model_to_ou1_data_global) -> p_value_ou1_vs_oum_global #0.003 

(((fit_ou1_model_to_ou1_data_remove_global$logLik-fit_oum_model_to_ou1_data_remove_global$logLik) < (remove_global_ou1_ouch@loglik - remove_global_oum_ouch@loglik)) %>% sum)/nrow(fit_ou1_model_to_ou1_data_remove_global) -> p_value_ou1_vs_oum_remove_global #0

(((fit_ou1_model_to_ou1_data_nearctic$logLik-fit_oum_model_to_ou1_data_nearctic$logLik) < (nearctic_ou1_ouch@loglik - nearctic_oum_ouch@loglik)) %>% sum)/nrow(fit_ou1_model_to_ou1_data_nearctic) -> p_value_ou1_vs_oum_nearctic #0.008 

(((fit_ou1_model_to_ou1_data_remove_nearctic$logLik-fit_oum_model_to_ou1_data_remove_nearctic$logLik) < (remove_nearctic_ou1_ouch@loglik - remove_nearctic_oum_ouch@loglik)) %>% sum)/nrow(fit_ou1_model_to_ou1_data_remove_nearctic) -> p_value_ou1_vs_oum_remove_nearctic #0

## Computing the power of a test that rejects the BM model with a false positive rate of 5%
## 1. Using the likelihood differences above for the BM and OU models fit to BM data, find the likelihood difference in the 5th percentile
## 2. Compute the likelihood difference between the BM model fit to the OU data and the OU model fit to the OU data for all simulated datasets
## 3. The number of likelihood differences computed in step #2 that are larger than the 5th percentile likelihood difference divided by the total number of simulated datasets gives an estimate of the power of the test
(((-2*(fit_bm_model_to_oum_data_global$logLik - fit_oum_model_to_oum_data_global$logLik)) > sort(-2*(fit_bm_model_to_bm_data_global$logLik-fit_oum_model_to_bm_data_global$logLik))[950]) %>% sum)/1000 -> power_to_reject_bm_global #1

(((-2*(fit_bm_model_to_oum_data_remove_global$logLik - fit_oum_model_to_oum_data_remove_global$logLik)) > sort(-2*(fit_bm_model_to_bm_data_remove_global$logLik-fit_oum_model_to_bm_data_remove_global$logLik))[950]) %>% sum)/1000 -> power_to_reject_bm_remove_global #0.999

(((-2*(fit_bm_model_to_oum_data_nearctic$logLik - fit_oum_model_to_oum_data_nearctic$logLik)) > sort(-2*(fit_bm_model_to_bm_data_nearctic$logLik-fit_oum_model_to_bm_data_nearctic$logLik))[950]) %>% sum)/1000 -> power_to_reject_bm_nearctic #1

(((-2*(fit_bm_model_to_oum_data_remove_nearctic$logLik - fit_oum_model_to_oum_data_remove_nearctic$logLik)) > sort(-2*(fit_bm_model_to_bm_data_remove_nearctic$logLik-fit_oum_model_to_bm_data_remove_nearctic$logLik))[950]) %>% sum)/1000 -> power_to_reject_bm_remove_nearctic #1


## Computing the power of a test that rejects the OU1 model with a false positive rate of 5%
## 1. Using the likelihood differences above for the OU1 and OUM models fit to OU1 data, find the likelihood difference in the 5th percentile
## 2. Compute the likelihood difference between the OU1 model fit to the OUM data and the OUM model fit to the OUM data for all simulated datasets
## 3. The number of likelihood differences computed in step #2 that are larger than the 5th percentile likelihood difference divided by the total number of simulated datasets gives an estimate of the power of the test
(((-2*(fit_ou1_model_to_oum_data_global$logLik - fit_oum_model_to_oum_data_global$logLik)) > sort(-2*(fit_ou1_model_to_ou1_data_global$logLik-fit_oum_model_to_ou1_data_global$logLik))[950]) %>% sum)/1000 -> power_to_reject_ou1_global # 0.988

(((-2*(fit_ou1_model_to_oum_data_remove_global$logLik - fit_oum_model_to_oum_data_remove_global$logLik)) > sort(-2*(fit_ou1_model_to_ou1_data_remove_global$logLik-fit_oum_model_to_ou1_data_remove_global$logLik))[950]) %>% sum)/1000 -> power_to_reject_ou1_remove_global #0.847

(((-2*(fit_ou1_model_to_oum_data_nearctic$logLik - fit_oum_model_to_oum_data_nearctic$logLik)) > sort(-2*(fit_ou1_model_to_ou1_data_nearctic$logLik-fit_oum_model_to_ou1_data_nearctic$logLik))[950]) %>% sum)/1000 -> power_to_reject_ou1_nearctic # 0.917

(((-2*(fit_ou1_model_to_oum_data_remove_nearctic$logLik - fit_oum_model_to_oum_data_remove_nearctic$logLik)) > sort(-2*(fit_ou1_model_to_ou1_data_remove_nearctic$logLik-fit_oum_model_to_ou1_data_remove_nearctic$logLik))[950]) %>% sum)/1000 -> power_to_reject_ou1_remove_nearctic #0.997
```


\newpage
Supplemental Figure 2. The likelihood ratio distributions for BM versus OU(3). 
```{r plot_bm_vs_oum, fig.height=4, fig.width=5, units='in', res=300, echo = FALSE}

## Plot the distributions of likelihood differences between BM and OUM for the four different datasets
par(mar=c(3,3,2,1), oma=rep(0,4), mfrow=c(2,2))
## Global
delta <- -2*(global_bm_ouch@loglik - global_oum_ouch@loglik)
delta.dist <- -2*(fit_bm_model_to_bm_data_global$logLik - fit_oum_model_to_bm_data_global$logLik)
delta.dist.2 <- -2*(fit_bm_model_to_oum_data_global$logLik - fit_oum_model_to_oum_data_global$logLik)
d1 <- density(delta.dist)
d2 <- density(delta.dist.2)
plot.new()
plot.window(xlim=range(c(d1$x,d2$x)), ylim=range(c(d1$y,d2$y)))
axis(1);axis(2);box('plot')
mtext(side=1, line=2, expression(delta))
mtext(side=2, line=2, 'Density')
mtext(side=3, line=0.25, "GMPD")
lines(d1$x, d1$y, col=gray(0.7))
with(d1, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.7, alpha=0.5)))
lines(d2$x, d2$y, col=gray(0.2))
with(d2, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.2, alpha=0.5)))
abline(v=delta, lwd=2, lty=2)
legend(x='topright', c('BM','OU3'), fill=c(gray(0.7),gray(0.2)), bty='n', cex=0.75)

delta <- -2*(remove_global_bm_ouch@loglik - remove_global_oum_ouch@loglik)
delta.dist <- -2*(fit_bm_model_to_bm_data_remove_global$logLik - fit_oum_model_to_bm_data_remove_global$logLik)
delta.dist.2 <- -2*(fit_bm_model_to_oum_data_remove_global$logLik - fit_oum_model_to_oum_data_remove_global$logLik)
d1 <- density(delta.dist)
d2 <- density(delta.dist.2)
plot.new()
plot.window(xlim=range(c(d1$x,d2$x)), ylim=range(c(d1$y,d2$y)))
axis(1);axis(2);box('plot')
mtext(side=1, line=2, expression(delta))
mtext(side=2, line=2, 'Density')
mtext(side=3, line=0.25, "GMPD (no single host)")
lines(d1$x, d1$y, col=gray(0.7))
with(d1, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.7, alpha=0.5)))
lines(d2$x, d2$y, col=gray(0.2))
with(d2, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.2, alpha=0.5)))
abline(v=delta, lwd=2, lty=2)

delta <- -2*(nearctic_bm_ouch@loglik - nearctic_oum_ouch@loglik)
delta.dist <- -2*(fit_bm_model_to_bm_data_nearctic$logLik - fit_oum_model_to_bm_data_nearctic$logLik)
delta.dist.2 <- -2*(fit_bm_model_to_oum_data_nearctic$logLik - fit_oum_model_to_oum_data_nearctic$logLik)
d1 <- density(delta.dist)
d2 <- density(delta.dist.2)
plot.new()
plot.window(xlim=range(c(d1$x,d2$x)), ylim=range(c(d1$y,d2$y)))
axis(1);axis(2);box('plot')
mtext(side=1, line=2, expression(delta))
mtext(side=2, line=2, 'Density')
mtext(side=3, line=0.25, "Nearctic")
lines(d1$x, d1$y, col=gray(0.7))
with(d1, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.7, alpha=0.5)))
lines(d2$x, d2$y, col=gray(0.2))
with(d2, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.2, alpha=0.5)))
abline(v=delta, lwd=2, lty=2)

delta <- -2*(remove_nearctic_bm_ouch@loglik - remove_nearctic_oum_ouch@loglik)
delta.dist <- -2*(fit_bm_model_to_bm_data_remove_nearctic$logLik - fit_oum_model_to_bm_data_remove_nearctic$logLik)
delta.dist.2 <- -2*(fit_bm_model_to_oum_data_remove_nearctic$logLik - fit_oum_model_to_oum_data_remove_nearctic$logLik)
d1 <- density(delta.dist)
d2 <- density(delta.dist.2)
plot.new()
plot.window(xlim=range(c(d1$x,d2$x)), ylim=range(c(d1$y,d2$y)))
axis(1);axis(2);box('plot')
mtext(side=1, line=2, expression(delta))
mtext(side=2, line=2, 'Density')
mtext(side=3, line=0.25, "Nearctic (no single host)")
lines(d1$x, d1$y, col=gray(0.7))
with(d1, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.7, alpha=0.5)))
lines(d2$x, d2$y, col=gray(0.2))
with(d2, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.2, alpha=0.5)))
abline(v=delta, lwd=2, lty=2)

```


\newpage
Supplemental Figure 3. The likelihood ratio distributions for OU(1) versus OU(3). 
```{r plot_ou1_vs_oum, fig.height=4, fig.width=5, units='in', res=300, echo = FALSE}

## Plot the distributions of likelihood differences between ou1 and OUM for the four different datasets
par(mar=c(3,3,2,1), oma=rep(0,4), mfrow=c(2,2))
## Global
delta <- -2*(global_ou1_ouch@loglik - global_oum_ouch@loglik)
delta.dist <- -2*(fit_ou1_model_to_ou1_data_global$logLik - fit_oum_model_to_ou1_data_global$logLik)
delta.dist.2 <- -2*(fit_ou1_model_to_oum_data_global$logLik - fit_oum_model_to_oum_data_global$logLik)
d1 <- density(delta.dist)
d2 <- density(delta.dist.2)
plot.new()
plot.window(xlim=range(c(d1$x,d2$x)), ylim=range(c(d1$y,d2$y)))
axis(1);axis(2);box('plot')
mtext(side=1, line=2, expression(delta))
mtext(side=2, line=2, 'Density')
mtext(side=3, line=0.25, "GMPD")
lines(d1$x, d1$y, col=gray(0.7))
with(d1, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.7, alpha=0.5)))
lines(d2$x, d2$y, col=gray(0.2))
with(d2, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.2, alpha=0.5)))
abline(v=delta, lwd=2, lty=2)
legend(x='topright', c('OU1','OU3'), fill=c(gray(0.7),gray(0.2)), bty='n', cex=0.75)

delta <- -2*(remove_global_ou1_ouch@loglik - remove_global_oum_ouch@loglik)
delta.dist <- -2*(fit_ou1_model_to_ou1_data_remove_global$logLik - fit_oum_model_to_ou1_data_remove_global$logLik)
delta.dist.2 <- -2*(fit_ou1_model_to_oum_data_remove_global$logLik - fit_oum_model_to_oum_data_remove_global$logLik)
d1 <- density(delta.dist)
d2 <- density(delta.dist.2)
plot.new()
plot.window(xlim=range(c(d1$x,d2$x)), ylim=range(c(d1$y,d2$y)))
axis(1);axis(2);box('plot')
mtext(side=1, line=2, expression(delta))
mtext(side=2, line=2, 'Density')
mtext(side=3, line=0.25, "GMPD (no single host)")
lines(d1$x, d1$y, col=gray(0.7))
with(d1, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.7, alpha=0.5)))
lines(d2$x, d2$y, col=gray(0.2))
with(d2, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.2, alpha=0.5)))
abline(v=delta, lwd=2, lty=2)

delta <- -2*(nearctic_ou1_ouch@loglik - nearctic_oum_ouch@loglik)
delta.dist <- -2*(fit_ou1_model_to_ou1_data_nearctic$logLik - fit_oum_model_to_ou1_data_nearctic$logLik)
delta.dist.2 <- -2*(fit_ou1_model_to_oum_data_nearctic$logLik - fit_oum_model_to_oum_data_nearctic$logLik)
d1 <- density(delta.dist)
d2 <- density(delta.dist.2)
plot.new()
plot.window(xlim=range(c(d1$x,d2$x)), ylim=range(c(d1$y,d2$y)))
axis(1);axis(2);box('plot')
mtext(side=1, line=2, expression(delta))
mtext(side=2, line=2, 'Density')
mtext(side=3, line=0.25, "Nearctic")
lines(d1$x, d1$y, col=gray(0.7))
with(d1, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.7, alpha=0.5)))
lines(d2$x, d2$y, col=gray(0.2))
with(d2, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.2, alpha=0.5)))
abline(v=delta, lwd=2, lty=2)

delta <- -2*(remove_nearctic_ou1_ouch@loglik - remove_nearctic_oum_ouch@loglik)
delta.dist <- -2*(fit_ou1_model_to_ou1_data_remove_nearctic$logLik - fit_oum_model_to_ou1_data_remove_nearctic$logLik)
delta.dist.2 <- -2*(fit_ou1_model_to_oum_data_remove_nearctic$logLik - fit_oum_model_to_oum_data_remove_nearctic$logLik)
d1 <- density(delta.dist)
d2 <- density(delta.dist.2)
plot.new()
plot.window(xlim=range(c(d1$x,d2$x)), ylim=range(c(d1$y,d2$y)))
axis(1);axis(2);box('plot')
mtext(side=1, line=2, expression(delta))
mtext(side=2, line=2, 'Density')
mtext(side=3, line=0.25, "Nearctic (no single host)")
lines(d1$x, d1$y, col=gray(0.7))
with(d1, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.7, alpha=0.5)))
lines(d2$x, d2$y, col=gray(0.2))
with(d2, polygon(x=c(min(x), x[x<max(x)], rev(x[x<max(x)]), min(x)), y=c(0, 0*x[x<max(x)], rev(y[x<max(x)]), 0), col=gray(0.2, alpha=0.5)))
abline(v=delta, lwd=2, lty=2)

```

\newpage
Supplemental Table 1. Summary table for evolutionary models of MPD in the Nearctic dataset with single host taxa = 0 
```{r echo = FALSE}
##with single host species MPD = 0
tibble(model=c("BM","OU1","OU3"),
       aicc=c(summary(nearctic_bm_ouch)$aic.c, 
              summary(nearctic_ou1_ouch)$aic.c,
              summary(nearctic_oum_ouch)$aic.c)
       ) %>% 
  mutate(deltaaicc=aicc-min(aicc), 
         pow=exp(-.5*deltaaicc),
         sumpow=sum(pow),
         waicc=pow/sumpow) %>%
  dplyr::select(c(1,2,3,6)) %>%
  arrange(desc(waicc)) -> nearctic_aic_OU
nearctic_aic_OU
```


Supplemental Table 2. Summary table for evolutionary models of MPD in the Nearctic dataset with single host taxa removed
```{r echo = FALSE}
##with single host species removed
tibble(model=c("BM","OU1","OU3"),
       aicc=c(summary(remove_nearctic_bm_ouch)$aic.c, 
              summary(remove_nearctic_ou1_ouch)$aic.c,
              summary(remove_nearctic_oum_ouch)$aic.c)
       ) %>% 
  mutate(deltaaicc=aicc-min(aicc), 
         pow=exp(-.5*deltaaicc),
         sumpow=sum(pow),
         waicc=pow/sumpow) %>%
  dplyr::select(c(1,2,3,6)) %>%
  arrange(desc(waicc)) -> remove_nearctic_aic_OU
remove_nearctic_aic_OU
```


Supplemental Table 3. Summary table for evolutionary models of MPD in the GMPD with single host taxa = 0 
```{r echo = FALSE}
##with single host species MPD = 0
tibble(model=c("BM","OU1","OU3"),
       aicc=c(summary(global_bm_ouch)$aic.c, 
              summary(global_ou1_ouch)$aic.c,
              summary(global_oum_ouch)$aic.c)
       ) %>% 
  mutate(deltaaicc=aicc-min(aicc), 
         pow=exp(-.5*deltaaicc),
         sumpow=sum(pow),
         waicc=pow/sumpow) %>%
  dplyr::select(c(1,2,3,6)) %>%
  arrange(desc(waicc)) -> global_aic_OU
global_aic_OU
```


Supplemental Table 4. Summary table for evolutionary models of MPD in the GMPD with single host taxa removed
```{r echo = FALSE}
tibble(model=c("BM","OU1","OU3"),
       aicc=c(summary(remove_global_bm_ouch)$aic.c, 
              summary(remove_global_ou1_ouch)$aic.c,
              summary(remove_global_oum_ouch)$aic.c)
       ) %>% 
  mutate(deltaaicc=aicc-min(aicc), 
         pow=exp(-.5*deltaaicc),
         sumpow=sum(pow),
         waicc=pow/sumpow) %>%
  dplyr::select(c(1,2,3,6)) %>%
  arrange(desc(waicc)) -> remove_global_aic_OU
remove_global_aic_OU
```




```{r adding_and_cleaning_host_and_parasite_trait_data, include = FALSE}


#### Load parasite body size data from Benesh, Lafferty, and Kuris 2017 ####
## Length and width measurments are in mm 
benesh_data_para<- read_csv("Benesh_et_al_2017/ecy1680-sup-0001-datas1/CLC_database_lifehistory.csv")
benesh_data_para$Host.species <-gsub(" ", "_", benesh_data_para$Host.species)

## remove parasites from Benesh et al. dataset that don't exist in the phylogeny
setdiff(benesh_data_para$Parasite.species, GMPDsmall$ParasiteCorrectedName)->nothanks
benesh_data_para[-which(benesh_data_para$Parasite.species %in% nothanks) ,] -> b_para_subset

## average the body size for rows that are the same species
subset(b_para_subset, Sex=="f") %>%
  group_by(Parasite.species) %>%
  summarise(mean_l=mean(Length, na.rm = TRUE)) ->data_fem_len

subset(b_para_subset, Sex=="f") %>%
  group_by(Parasite.species) %>%
  summarise(mean_w=mean(Width, na.rm = TRUE)) ->data_fem_wid

## combine length and width to calculate 'body size' (L*W)?
b.size<-left_join(data_fem_len, data_fem_wid, by="Parasite.species")
b.size$ratio<-(b.size$mean_l*b.size$mean_w)

write.csv(b.size, "bodysizepara.csv")


#### Trim the tree to include only those parasite species for which there is body size data ####

##nearctic mpd, single host = 0 
setdiff(nearctic_tree$tip.label, b.size$Parasite.species)->no_body_size 
nearctic_pgls <- drop.tip(nearctic_tree, nearctic_tree$tip.label[which(nearctic_tree$tip.label%in%no_body_size)])

# make sure body size is in the same order as the parasite phylogeny
nearctic.b.size <- b.size[match(nearctic_pgls$tip.label, b.size$Parasite.species),]


##nearctic mpd, single host removed 
setdiff(remove_nearctic_tree$tip.label, b.size$Parasite.species)->no_body_size1 
remove_nearctic_pgls <- drop.tip(remove_nearctic_tree, remove_nearctic_tree$tip.label[which(remove_nearctic_tree$tip.label%in%no_body_size1)])

# make sure body size is in the same order as the parasite phylogeny
remove.nearctic.b.size <- b.size[match(remove_nearctic_pgls$tip.label, b.size$Parasite.species),]


##global mpd, single host = 0 
setdiff(global_tree$tip.label, b.size$Parasite.species)->no_body_size2 
global_pgls <- drop.tip(global_tree, global_tree$tip.label[which(global_tree$tip.label%in%no_body_size2)])

# make sure body size is in the same order as the parasite phylogeny
global.b.size <- b.size[match(global_pgls$tip.label, b.size$Parasite.species),]

##global mpd, single host removed
setdiff(remove_global_tree$tip.label, b.size$Parasite.species)->no_body_size3 
remove_global_pgls <- drop.tip(remove_global_tree, remove_global_tree$tip.label[which(remove_global_tree$tip.label%in%no_body_size3)])

# make sure body size is in the same order as the parasite phylogeny
remove.global.b.size <- b.size[match(remove_global_pgls$tip.label, b.size$Parasite.species),]



#### adding in host trait data (mass.g) from PHYLACINE ####
#Faurby, S., Davis, M., Pedersen, R. ., Schowanek, S. D., Antonelli, A., & Svenning, J.C. (In Press). PHYLACINE 1.2: The Phylogenetic Atlas of Mammal Macroecology. Ecology
#load mammal data
Trait_data <- read_csv("Trait_data.csv")

#correct for spaces in the binomial names
Trait_data$Binomial.1.2<-gsub("_", " ", Trait_data$Binomial.1.2)

#make column called "Host" in Trait_data
Trait_data$HostCorrectedName<-Trait_data$Binomial.1.2

#merge 
left_join(GMPDsmall, Trait_data, by = "HostCorrectedName")->ss
host_trait_data<-data.frame("Parasite" = ss$ParasiteCorrectedName,
                            "Host" = ss$HostCorrectedName, 
                            "Mass.g" = ss$Mass.g)


host_trait_data %>% 
  group_by(Parasite) %>% 
  summarise(avgMass=mean(Mass.g,na.rm=T)) -> Host_mass


##subset host data for each group

#nearctic with single hosts = 0
setdiff(Host_mass$Parasite, nearctic_pgls$tip.label)-> getout
nearctic_traits<-Host_mass[-which(Host_mass$Parasite %in% getout),]
#put in tree tip label order
nearctic_traits <- nearctic_traits[match(nearctic_pgls$tip.label, nearctic_traits$Parasite),]


#nearctic with single hosts removed
setdiff(Host_mass$Parasite, remove_nearctic_pgls$tip.label)-> getout1
remove_nearctic_traits<-Host_mass[-which(Host_mass$Parasite %in% getout1),]
#put in tree tip label order
remove_nearctic_traits <- remove_nearctic_traits[match(remove_nearctic_pgls$tip.label, remove_nearctic_traits$Parasite),]


#global with single hosts = 0
global_traits<-Host_mass
#put in tree tip label order
global_traits <- global_traits[match(global_pgls$tip.label, global_traits$Parasite),]


#global with single hosts removed
setdiff(Host_mass$Parasite, remove_global_pgls$tip.label)-> getout2
remove_global_traits<-Host_mass[-which(Host_mass$Parasite %in% getout2),]
#put in tree tip label order
remove_global_traits <- remove_global_traits[match(remove_global_pgls$tip.label, remove_global_traits$Parasite),]


#### clean parasite transmission data ####

#nearctic with single hosts = 0
nearctic_trans<-GMPDparasite[which(GMPDparasite$CorrectName %in% nearctic_pgls$tip.label),]
nearctic_trans<-nearctic_trans[match(nearctic_pgls$tip.label, nearctic_trans$CorrectName),]

#neartic with single species removed
remove_nearctic_trans<-GMPDparasite[which(GMPDparasite$CorrectName %in% remove_nearctic_pgls$tip.label),]
remove_nearctic_trans<-remove_nearctic_trans[match(remove_nearctic_pgls$tip.label, remove_nearctic_trans$CorrectName),]


#global with single hosts = 0
global_trans<-GMPDparasite[which(GMPDparasite$CorrectName %in% global_pgls$tip.label),]
global_trans<-global_trans[match(global_pgls$tip.label, global_trans$CorrectName),]

#global with single species removed
remove_global_trans<-GMPDparasite[which(GMPDparasite$CorrectName %in% remove_global_pgls$tip.label),]
remove_global_trans<-remove_global_trans[match(remove_global_pgls$tip.label, remove_global_trans$CorrectName),]

###NEARCTIC DATASET####

##with single host species MPD = 0

#clean nearctic.mam.mpd
pgls.nearctic.para.mpd<-nearctic.para.mpd[match(nearctic_pgls$tip.label, names(nearctic.para.mpd))]

nearctic_dat<-data.frame(taxa=nearctic_pgls$tip.label,
                para.mpd=pgls.nearctic.para.mpd,
                para.l=nearctic.b.size$mean_l,
                para.w=nearctic.b.size$mean_w,
                para.b=nearctic.b.size$ratio,
                close=as.factor(nearctic_trans$close),
                nonclose=as.factor(nearctic_trans$nonclose),
                intermediate=as.factor(nearctic_trans$intermediate),
                host_mass=nearctic_traits$avgMass)

## constructs the data.frame of the taxa names and data in the order of the tree tips
nearctic_cdat<-comparative.data(data=nearctic_dat, phy=nearctic_pgls, names.col = "taxa")
print(nearctic_cdat)

## model testing ##

#all predictors
nearctic_mod <- pgls(para.mpd~ para.l + close + nonclose + intermediate + host_mass, nearctic_cdat, lambda = .384)
summary(nearctic_mod)
anova(nearctic_mod)

#removed parasite length 
nearctic_mod2<-pgls(para.mpd ~ close + nonclose + intermediate + host_mass, nearctic_cdat, lambda=.384) #host mass is significant
summary(nearctic_mod2)
anova(nearctic_mod2)

#removed parasite length and close transmission
nearctic_mod3<-pgls(para.mpd ~ nonclose + intermediate + host_mass, nearctic_cdat, lambda=.384) 
summary(nearctic_mod3)
anova(nearctic_mod3)

anova(nearctic_mod2, nearctic_mod3)

#removed parasite length and close transmission + intermediate ##### BEST MODEL #####
nearctic_mod4<-pgls(para.mpd ~ nonclose + host_mass, nearctic_cdat, lambda=.384) 
summary(nearctic_mod4)
anova(nearctic_mod4)

anova(nearctic_mod3, nearctic_mod4)

```


Best model fit with PGLS for the Nearctic helminth phylogeny, with single host taxa included.
```{r echo = FALSE}
#only host mass ##### BEST MODEL #####
nearctic_mod5<-pgls(para.mpd ~  host_mass, nearctic_cdat, lambda=.384) 
summary(nearctic_mod5)
#anova(nearctic_mod5)

```



```{r pgls_nearctic_shR, include=FALSE}
###NEARCTIC DATASET####

##with single host species removed

#clean nearctic.mam.mpd
pgls.remove.n.para.mpd<-nearctic.para.mpd[match(remove_nearctic_pgls$tip.label, names(nearctic.para.mpd.remove))]

remove_nearctic_dat<-data.frame(taxa=remove_nearctic_pgls$tip.label,
                para.mpd=pgls.remove.n.para.mpd,
                para.l=remove.nearctic.b.size$mean_l,
                para.w=remove.nearctic.b.size$mean_w,
                para.b=remove.nearctic.b.size$ratio,
                close=as.factor(remove_nearctic_trans$close),
                nonclose=as.factor(remove_nearctic_trans$nonclose),
                intermediate=as.factor(remove_nearctic_trans$intermediate),
                host_mass=remove_nearctic_traits$avgMass)

## constructs the data.frame of the taxa names and data in the order of the tree tips
remove_nearctic_cdat<-comparative.data(data=remove_nearctic_dat, phy=remove_nearctic_pgls, names.col = "taxa")
print(remove_nearctic_cdat)

## model testing ##

#all predictors 
remove_nearctic_mod <- pgls(para.mpd~ para.l + close + nonclose + intermediate + host_mass, remove_nearctic_cdat, lambda = .926)
summary(remove_nearctic_mod)
anova(remove_nearctic_mod)

#removed host mass  ####### BEST MODEL #######
remove_nearctic_mod2<-pgls(para.mpd~ para.l + close + nonclose + intermediate, remove_nearctic_cdat, lambda = .926)
summary(remove_nearctic_mod2) 
anova(remove_nearctic_mod2)

anova(remove_nearctic_mod, remove_nearctic_mod2)


#removed host mass and close transmission
remove_nearctic_mod3<-pgls(para.mpd~ para.l + nonclose + intermediate , remove_nearctic_cdat, lambda = .926)
summary(remove_nearctic_mod3) #worse model than ^^^
anova(remove_nearctic_mod3)

```

Best model fit with PGLS for the Nearctic helminth phylogeny, without single host taxa.
```{r echo=FALSE}
remove_nearctic_mod2<-pgls(para.mpd~ para.l + close + nonclose + intermediate, remove_nearctic_cdat, lambda = .926)
summary(remove_nearctic_mod2) 
anova(remove_nearctic_mod2)
```



```{r pgls_global_sh0, include=FALSE}

###GLOBAL DATASET####

##with single host species MPD = 0

#clean global.mam.mpd
pgls.global.para.mpd<-global.para.mpd[match(global_pgls$tip.label, names(global.para.mpd))]

global_dat<-data.frame(taxa=global_pgls$tip.label,
                para.mpd=pgls.global.para.mpd,
                para.l=global.b.size$mean_l,
                para.w=global.b.size$mean_w,
                para.b=global.b.size$ratio,
                close=as.factor(global_trans$close),
                nonclose=as.factor(global_trans$nonclose),
                intermediate=as.factor(global_trans$intermediate),
                host_mass=global_traits$avgMass)

## constructs the data.frame of the taxa names and data in the order of the tree tips
global_cdat<-comparative.data(data=global_dat, phy=global_pgls, names.col = "taxa")
print(global_cdat)

## model testing ##

#all predictors
global_mod <- pgls(para.mpd~ para.l + close + nonclose + intermediate + host_mass, global_cdat, lambda = .607)
summary(global_mod)
anova(global_mod)

#removed host mass
global_mod2 <- pgls(para.mpd~ para.l + close + nonclose + intermediate, global_cdat, lambda = .607)
summary(global_mod2)
anova(global_mod2)


#removed nonclose and host mass
global_mod3 <- pgls(para.mpd~ para.l + close + intermediate, global_cdat, lambda = .607)
summary(global_mod3) 
anova(global_mod3)

#only parasite length and close transmission
global_mod4<-pgls(para.mpd ~ para.l + close, global_cdat, lambda = .607)
summary(global_mod4) 
anova(global_mod4)

#only close transmission  #### BEST MODEL ####
global_mod5<-pgls(para.mpd ~ close, global_cdat, lambda = .607)
summary(global_mod5) 
anova(global_mod5)

anova(global_mod4, global_mod5)


```


Best model fit with PGLS for the GMPD helminth phylogeny, with single host taxa included.
```{r echo=FALSE}
#only close transmission  #### BEST MODEL ####
global_mod5<-pgls(para.mpd ~ close, global_cdat, lambda = .607)
summary(global_mod5) 
anova(global_mod5)
```


```{r include= FALSE}
###GLOBAL DATASET####

##with single host species removed

#clean nearctic.mam.mpd
pgls.remove.g.para.mpd<-global.para.mpd[match(remove_global_pgls$tip.label, names(global.para.mpd.remove))]

remove_global_dat<-data.frame(taxa=remove_global_pgls$tip.label,
                para.mpd=pgls.remove.g.para.mpd,
                para.l=remove.global.b.size$mean_l,
                para.w=remove.global.b.size$mean_w,
                para.b=remove.global.b.size$ratio,
                close=as.factor(remove_global_trans$close),
                nonclose=as.factor(remove_global_trans$nonclose),
                intermediate=as.factor(remove_global_trans$intermediate),
                host_mass=remove_global_traits$avgMass)

## constructs the data.frame of the taxa names and data in the order of the tree tips
remove_global_cdat<-comparative.data(data=remove_global_dat, phy=remove_global_pgls, names.col = "taxa")
print(remove_global_cdat)

## model testing ##

#all predictors
remove_global_mod <- pgls(para.mpd~ para.l + close + nonclose + intermediate + host_mass, remove_global_cdat, lambda = .895)
summary(remove_global_mod)
anova(remove_global_mod)

#removed host mass
remove_global_mod2 <- pgls(para.mpd~ para.l + close + nonclose + intermediate, remove_global_cdat, lambda = .895)
summary(remove_global_mod2) 
anova(remove_global_mod2)


#removed nonclose and host mass
remove_global_mod3 <- pgls(para.mpd~ para.l + close + intermediate, remove_global_cdat, lambda = .895)
summary(remove_global_mod3) 
anova(remove_global_mod3)


#just parasite length + close
remove_global_mod4 <- pgls(para.mpd~ para.l  + close, remove_global_cdat, lambda = .895)
summary(remove_global_mod4) 
anova(remove_global_mod4)

#just close transmission  #### BEST MODEL ####
remove_global_mod5 <- pgls(para.mpd~ close, remove_global_cdat, lambda = .895)
summary(remove_global_mod5) 
anova(remove_global_mod5)

```


Best model fit with PGLS for the GMPD helminth phylogeny, without single host taxa.
```{r echo=FALSE}
#just close transmission  #### BEST MODEL ####
remove_global_mod5 <- pgls(para.mpd~ close, remove_global_cdat, lambda = .895)
summary(remove_global_mod5) 
anova(remove_global_mod5)
```

